---
output: html_document
---
```{r}
library(datasets)
library(plyr)
library(dplyr)
library(reshape2)
library(knitr)
library(ggplot2)
```

### Title: PeerAssessment-2  
Synopsis:
  This report is generated for the 2nd assignment of Reproducible Research Course.
  It involves exploring the NOAA database. This database tracks characteristics of
  major storms and weather events in the United States, including when and where they occur, 
  as well as estimates of any fatalities, injuries, and property damage. 
  For this report we will analyse the events that cause maximum number of
  bodily injuries and human fatalities and, have the maximum financial consequences.


### Data Processing:
To start with, we download and read storm data into r. Then create the working subset.

```{r, cache=TRUE} 
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2", 
 dest="stormData",method = "wininet")
 
stD <- read.csv("stormData", sep = "," )
substD <- subset(stD, FATALITIES > 0 & INJURIES > 0, select=c(STATE,INJURIES,FATALITIES,EVTYPE))

```
Based on the information in the "stormData", this Report attempts to answer following questions:
 
         1.Across the United States, which types of events (as indicated in the EVTYPE variable) 
           are most harmful with respect to population health?
         2.Across the United States, which types of events have the greatest economic 
           consequences?

To answer the first questions:
 
1. We select both "INJURIES" and "FATALITIES" to identify the EVTYPE that are most damaging  
   to human health. 
   
```{r}
hlth_dmg <- aggregate(INJURIES + FATALITIES ~ EVTYPE + STATE, data=substD,max)
```
2. Assign more relevant names to the result set.
```{r}
xnames <- c("EVENT", "STATE", "HEALTH_DMG" )
names(hlth_dmg) <- xnames
```

3. Process the weather events that are most damaging to population health across all states.
  
```{r} 

max_dmg <- subset(hlth_dmg, HEALTH_DMG >= 1 & HEALTH_DMG <= max(HEALTH_DMG), select = ordered(c(EVENT,STATE,HEALTH_DMG )))
xmaxD <- head(max_dmg[with(max_dmg, order(desc(HEALTH_DMG), STATE)), ],52)
```
4.  We report the results of our analysis in the results section. We also plot the results
    in the plots sections. 
 



To answer the second question

1. We select both "CROPDMG" and "PROPDMG" to identify the EVTYPE that have the greatest economic consequences.

```{r, cache=TRUE}   
substE <- subset(stD, CROPDMG > 0 & PROPDMG > 0, select=c(STATE,CROPDMG,PROPDMG,EVTYPE))
eco_dmg <- aggregate(CROPDMG + PROPDMG ~ EVTYPE + STATE, data=substE,max)
```
2. Assign more relevant names to the result set.
```{r}

enames <- c("EVENT", "STATE", "DAMAGE")
names(eco_dmg) <- enames
```
3. Process the weather events responsible for greatest economic consequences across all states in
  the United States. 
  
```{r}

max_edmg <- subset(eco_dmg, DAMAGE >= 1 & DAMAGE <= max(DAMAGE), select = ordered(c(EVENT,STATE,DAMAGE )))
xmaxE <- head(max_edmg[with(max_edmg, order(desc(DAMAGE),STATE)), ],52)
```
  
4. We report the results of our analysis in the results section. We also plot the results in the
  plots sections.


### Results:
our analysis show:

1.   TORNADO is the most damaging weather event that causes bodily injuries and fatalities
     to population across all states. It also shows that Texas has the highest count of
     injuries and fatalities from this event.

```{r}
xmaxD
```
2.   HIGH WIND is the most damaging weather event responsible for maximum economic impact
     across all states. It also shows that NC incurs the highest economic impact from this event.

```{r}
xmaxE
```

### Plots

```{r, plot1 }
 
qplot(HEALTH_DMG,EVENT, data=xmaxD,facets=.~ STATE, aes(group = 2, geom_line()), 
main="Health damage to population accross U.S",
ylab="Weather Events",
xlab="Maximum damage by weather related events")
```
```{r}

```{r, plot2 }
 
qplot(DAMAGE,EVENT, data=xmaxE,facets=.~ STATE, aes(group = 2, geom_line()), 
main="Economic consequences across the U.S",
ylab="Weather Events",
xlab="Maximum damage by weather related events")
```



 
  


